---
title: "Section 2: summarizing single variables"
subtitle: "<br/>STOR 155.02, Spring '21"
date: "updated `r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      highlightStyle: solarized-light
      highlightLines: true
      countIncrementalSlides: false
---
class: left, middle

```{r xaringan-themer, include=FALSE, warning=FALSE}
#https://pkg.garrickadenbuie.com/xaringanthemer/articles/themes.html
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = F, warning = F, message = F)
library(xaringanthemer)
library(showtext)

style_solarized_light()
# style_mono_accent(
#   base_color = "#1c5253",
#   header_font_google = google_font("Josefin Sans"),
#   text_font_google   = google_font("Montserrat", "300", "300i"),
#   code_font_google   = google_font("Fira Mono")
# )
```

## What you will learn

- xxx sample distributions
- xxx plots: bar graphs, histograms
- xxx centrality statistics: mean, median
- xxx spread statistics: range, quantiles, variance and standard deviation

## Resources

- xxx textbook

---
class: inverse, center, middle

# Looking at data

---
class: left, top
## Example data for today: Duke



---
class: left, top

## Definition: sample distribution
>> The sample distribution of a variable tells us what values it takes and how often it takes these values in our dataset.


#### this is about the values a variable actually has in your data

#### does *not* tell you about all possible values your variable could have if you got more or different data

---
class: left, top
## Looking at sample distributions: depends on the variable type

.pull-left[
### Categorical:

#### bar graphs

#### pie charts
]


.pull-right[
### Numeric/quantitative:

#### histograms
]

---
class: left, top
## Looking at categorical data

### frequency statistics

- **frequency** of a category is the **number of observations** in that category
- **relative frequency** is number of **observations in a category / total** number of observations 

### bar graph

- bar height represents frequency or relative frequency for each category
- the frequency must first be computed

### pie chart

- area of a pie slice represents the relative frequency of a category
- these can be **hard to read** because areas are difficult to estimate by eye
- bar charts preferred usually

---
class: left, top
## Example: Fake news on facebook

.pull-left[
### dataset

- data on Facebook popularity of posts by fake news websites in 2018
- identified and collected by Buzzfeed News
- [Biggest fake news hits on Facebook in 2018](https://www.buzzfeednews.com/article/craigsilverman/facebook-fake-news-hits-2018)
- source: [Buzzfeed github site](https://github.com/BuzzFeedNews/2018-12-fake-news-top-50)
- dataset on our course github page as well
]


.pull-right[
### observations and variables

- each row is represents an article
- `fb_engagement` is essentially number of likes post received
- `category` is a buzzfeed-created categorical variable
]

---
class: left, top


### Bar chart of observations, select categories
```{r}
library(readr)
library(dplyr)
library(ggplot2)
fb <- read_csv("https://raw.githubusercontent.com/brendanrbrown/stor155_sp21/main/data/fakebook_buzzfeed_2018.csv")
```

```{r, dpi=300, fig.heigh=10, fig.width=12}
filter(fb, category %in% c("Obama Spy", "Crime", "Medical", "Business", "Sports", "Music")) %>%
  ggplot(aes(x = category)) + geom_bar(fill = '#dc322f') + theme_bw() + ggtitle("Frequency of fake news postings on Facebook", subtitle = "Buzzfeed news, 2018")
```

